Efficiently stable presentations from error-correcting codes

### 2023-10-02 efficient_stability.tex
Updates from Overleaf

	Let $B(\ell_2)$ be the von Neumann algebra of bounded operators on $\ell_2$, the Hilbert space of convergent sequences in $\C^\Z$ equipped with the usual Euclidean norm (for which we let $(e_i)_{i \in \Z}$ denote the standard basis). We denote $\mM_\infty = \mM \overline{\otimes} B(\ell_2)$, where the overline denotes closure for the operator topology. $\mM_\infty$ is a von Neumann algebra equipped with the (infinite) trace $\tau_\infty = \tau \otimes \Tr$, with $\Tr(x)=\sum_{i\in \Z} e_i^T X e_i$ the trace on $B(\ell_2)$. We generally identify $\mM$ with the ``corner'' $\mM\otimes e_{1,1}\subset \mM_\infty$.

@@

\begin{proof}
If $\eps\geq \frac{1}{2}$ the conclusion is trivial, so assume $\eps<\frac{1}{2}$. 
Define 
\[\tilde{Q}_a = w^* T_a w  + \frac{1}{|\mA|}\big(\Id_\mM - w^* w\big) \in \mM\;.\]
Then $\{\tilde{Q}_a\}$ is a POVM on $\mM$. Moreover, 
\begin{align*}
\sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) &\geq \sum_a \tau^\mM \big( \big(w^* T_a w \big)^2 \big) \\
&= \sum_a \tau^\mM \big(  w^* T_a w w^*T_a w \big)\\
&= \sum_a \tau^\mM \big(  w^* T_a  P T_a w \big) - \sum_a \tau^\mM \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \sum_a \tau_\infty \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \tau_\infty\Big(\big( P - w w^*\big)\Big(\sum_a  T_a w w^* T_a\Big)\Big)\\ 
&\geq 1- \eps- \tau_\infty\big( P- w w^*\big)\;,
\end{align*}
where the third line uses that $T_aPT_a=T_a$, $\sum_a T_a = \Id_\mN$ and the definition of $\eps$ for the first term, and for the second the fact that for $A\in\mM$, $\tau^\mM(A)=\tau_\infty(A)$ by definition of $\tau_\infty$ and the identification of $\mM$ with a ``corner'' in $\mM_\infty$, the fourth line uses cyclicity of the trace for the second, and the last uses $\|ww^*\|,\|\sum_a T_a\|_\infty\leq 1$. By assumption, 
\begin{align}
\tau_\infty\big( P- w w^*\big) \,\leq\, \eps\, \tau_\infty(P)\,\leq \frac{\eps}{1-\eps}\;,\label{eq:bound_on_P-ww*_in_trace_infty}
\end{align}
where the last inequality is because by definition, $\tau^N(P)=1$, thus
\[1-\eps \,\leq\, \tau^\mN(ww^*) \,=\, \frac{\tau_\infty(ww^*)}{\tau_\infty(P)}\,=\, \frac{\tau_\infty(w^*w)}{\tau_\infty(P)}  \,\leq\, \frac{1}{\tau_\infty(P)}\]
since $\tau_\infty(w^* w) = \tau^\mM(w^* w)$ and $w^*w\leq I_\mM$. Overall, 
\[ \sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) \,\geq\, 1-\eps-\frac{\eps}{1-\eps}\,\geq\, 1-3\eps\;.\]
To conclude we apply~\cite[Theorem 1.2]{de2021orthogonalization} to obtain a projective measurement $\{Q_a\}$ on $\mM$ such that 
\begin{equation*}
\sum_a \big\|{Q}_a - \tilde{Q}_a \big\|^2_\tau \,=\, 27\eps\;.
\end{equation*}
Finally,
\begin{align*}
\sum_a \big\|{Q}_a - w^*{T}_a w\big\|^2_\tau &= \sum_a \Big\|{Q}_a - \tilde{Q}_a  + \frac{1}{|\mA|}\big(I_\mM - w^* w\big) \Big\|^2_2\\
&\leq  \sum_a 2\big\|{Q}_a - \tilde{Q}_a\big\|_\tau^2  + 2\frac{1}{|\mA|}\big\|I_\mM - w^* w\big\|_\tau^2 \\
&\leq 54 \eps + 2 \tau^\mM( (I_\mM - w^* w)^2 ) \\
&\leq 54 \eps + 2 \tau^\mM(I_\mM - w^* w ) \\
&\leq 56 \eps\;,
\end{align*}
where the second line is by the triangle inequality, the fourth line is due to the fact that $I_\mM - w^* w$ is positive and has operator norm at most $1$, and the last line is by $\tau^\mM(I_\mM - w^* w ) \leq \eps$.
\end{proof}

@@

As one would naturally expect, the dependence of $\delta_k$ on $k$ depends on both the diameter of $\Z_2^k$ for the presentation~\eqref{eq:z2-efficient}, and its \emph{Dehn function} --- namely, the minimal volume of a Van Kampen diagram with perimeters of length $3$. By expressing each element of $\Z_2^k$ as a product of generators in the natural way, applying Corollary~\ref{cor:lin-test} it is possible to show that $\delta_k=O(k^2\eps)$ in Lemma~\ref{lem:eff-z2}. With more work one can get $\delta_k=O(k\eps)$, see~\cite[Theorem 3.2]{chao2017overlapping}. This turns up to be  tight:
\begin{claim}
    Let $\mu_R$ be the uniform distribution over the commutation relations in the presentation $\langle S\colon R\rangle$ from \eqref{eq:z2-efficient}, and $\mu_S$ the marginal distribution from $\mu_R$, which is the uniform one on $S=\{x_1,...,x_k\}$. There is a $(\nicefrac{2}{\binom{k}{2}},\mu_R)$-almost homomorphism of $\langle S\colon R\rangle$ which is at least $\nicefrac{c}{k}$ far from any homomorphism. Namely, the modulus of stability of \eqref{eq:z2-efficient} is $\Omega(k\eps)$.
\end{claim}

@@

Let $\{B_i\}_{i=1}^k$ be a collection of unitaries in $(\mN,\tau^\mN)$ --- where $\mN$ is $P\mM_\infty P$  for some projection $P$ of finite trace and $\tau^\mN=\frac{\tau_\infty}{\tau_\infty(P)}$ --- which satisfy all the relations in $R$. Namely, each $B_i$ is an order $2$ unitary and they all commute. For every $x\in \Z_2^k$ let  $B_x=\prod_{i=1}^kB_i^{x_i}$. Now, we can apply on this representation of $\Z_2^k$ the same operation as we did in the proof of Corollary \ref{cor:lin-test}: For every $u\in \Z_2^k$ we define $Q_u=\Expectation_x(-1)^{u\cdot x}B_X$. Then, again, $\{Q_u\}$ is a projective measurement on $\mN$. By Lemma \ref{lem:pull-back}, for every partial isometry $w\in P\mM_\infty \Id_\mM$ we have a projective measurement $\{P_u\}_{u\in \Z_2^k}$  on $\mM=M_{2^{k+1}}(\complex)$ such that 
\[
\sum_{u\in \Z_2^k}\Vert P_u -w^*Q_uw\Vert_\tau^2\leq 56\theta,
\]
where $\theta=\max\{\tau(\Id_\mM-w^*w),\tau'(P-ww^*)\}$.
Let $C_x=\sum_{u\in \Z_2^k}(-1)^{u\cdot x}P_u$ for every $x\in \Z_2^k$. Then, as in \eqref{eq:lin-test-2}, we have
\[
\begin{split}
 \Expectation_{x\in \Z_2^k}\Vert C_x-w^*B_xw\Vert_\tau^2\leq 56\theta.
\end{split}
\]
But, for every $x,y\in \Z_2^k$, we have
\begin{align}
      \Vert C_x-w^*B_xw\Vert_\tau&=\Vert C_yC_{x+y}-w^*B_yB_{x+y}w\Vert_\tau \notag\\
    &=\Vert C_yC_{x+y}-C_yw^*B_{x+y}w+C_yw^*B_{x+y}w-w^*B_yww^*B_{x+y}w \notag\\ &+w^*B_yww^*B_{x+y}w-w^*B_yB_{x+y}w\Vert_\tau\notag \\
    &\leq \Vert C_{x+y}-w^*B_{x+y}w\Vert_\tau+\Vert C_y-w^*B_yw\Vert_\tau+\Vert w^*B_y(ww^*-P)B_{x+y}w\Vert_{\tau}.\label{eq:lower_bound_stab_standard_pres}
\end{align}
Now, 
\begin{align}
    \Vert w^*B_y(ww^*-P)B_{x+y}w\Vert_{\tau}^2&=\tau^\mM(w^*B_{x+y}(ww^*-P)B_yww^*B_y(ww^*-P)B_{x+y}w)\notag\\
    &\leq \tau^\infty(w^*B_{x+y}(ww^*-P)B_yww^*B_y(ww^*-P)B_{x+y}w)\notag\\
     &\leq \tau^\infty((ww^*-P)B_yww^*B_y(ww^*-P)B_{x+y}ww^*B_{x+y})\notag\\
     &\leq \tau^\infty((ww^*-P)B_yww^*B_y(ww^*-P))\notag\\
     &\leq \tau^\infty(\underbrace{(ww^*-P)^2}_{P-ww^*}B_yww^*B_y)\notag\\
     &\leq \frac{\theta}{1-\theta}.\label{eq:lower_bound_stab_standard_pres1}
\end{align}
where, the first line is by definition, the second line is by the fact $\mM$ is a corner and the evaluated operator is positive, the third and fifth lines are by the cyclicity of the trace, the fourth and sixth lines use von Neumann's trace inequality and the fact $B_{x+y}ww^*B_{x+y}$ and $B_{y}ww^*B_{y}$ are contractions, as well as \eqref{eq:bound_on_P-ww*_in_trace_infty}.
Combining \eqref{eq:lower_bound_stab_standard_pres} and \eqref{eq:lower_bound_stab_standard_pres1}, we deduce that 
\begin{align}
    \Vert C_x-w^*B_xw\Vert_\tau^2&=\Expectation_{y}\Vert C_x-w^*B_xw\Vert_\tau^2\notag\\
    &\leq 3\Expectation_{y} \Vert C_{x+y}-w^*B_{x+y}w\Vert_\tau^2+\Vert C_y-w^*B_yw\Vert_\tau^2+\Vert w^*B_y(ww^*-P)B_{x+y}w\Vert_{\tau}^2\notag\\
    &\leq 3(56\theta +56\theta +\frac{\theta}{1-\theta}).
\end{align}
But, by denoting $C_j=C_{e_j}$, we have
\begin{align*}
       \sqrt{2}=\Vert A_1A_2A_1A_2-\underbrace{\Id_\mM}_{C_1C_2C_1C_2}\Vert_\tau
       \leq 2\Vert A_1-C_1\Vert_\tau+2\Vert A_2-C_2\Vert_\tau,
\end{align*}
where the inequality uses the fact the $A_i$'s and $C_i$'s are unitaries together with the triangle inequality, and the equality is a straightforward calculation of their action on the computational basis elements --- which are bit strings of length $k+1$. Hence,
\[
\Expectation \Vert A_i-C_i\Vert_\tau ^2\geq \frac{1}{k}(\Vert A_1-C_1\Vert_\tau^2+\Vert A_2-C_2\Vert_\tau^2)\geq \frac{1}{4k}.
\]
Combining everything, we get
\begin{align*}
    \Expectation \Vert A_i-w^*B_iw\Vert_\tau ^2&\geq \Expectation \Vert A_i-C_i\Vert_\tau ^2-\Expectation \Vert C_i-w^*B_iw\Vert_\tau ^2\\
    &\geq \frac{1}{4k}-336\theta-\frac{3\theta}{1-\theta}.
\end{align*}
\end{proof}

>>>>

	Let $B(\ell_2)$ be the von Neumann algebra of bounded operators on $\ell_2$, the Hilbert space of convergent sequences in $\C^\Z$ equipped with the usual Euclidean norm (for which we let $(e_i)_{i \in \Z}$ denote the standard basis). We denote $\mM_\infty = \mM \overline{\otimes} B(\ell_2)$, where the overline denotes closure for the operator topology. $\mM_\infty$ is a von Neumann algebra equipped with the (infinite) trace $\tau_\infty = \tau \otimes \Tr$, with $\Tr(X)=\sum_{i\in \Z} e_i^T X e_i$ the trace on $B(\ell_2)$. We generally identify $\mM$ with the ``corner'' $\mM\otimes \Id_{1}\subset \mM_\infty$, where $\Id_1$ is the projection on the $1^{\rm st}$ coordinate in $\complex^\Z$.

@@

\begin{proof}
If $\eps\geq \frac{1}{2}$ the conclusion is trivial, so assume $\eps<\frac{1}{2}$. 
Define 
\[\tilde{Q}_a = w^* T_a w  + \frac{1}{|\mA|}\big(\Id_\mM - w^* w\big) \in \mM\;.\]
Then $\{\tilde{Q}_a\}$ is a POVM on $\mM$. Moreover, 
\begin{align*}
\sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) &\geq \sum_a \tau^\mM \big( \big(w^* T_a w \big)^2 \big) \\
&= \sum_a \tau^\mM \big(  w^* T_a w w^*T_a w \big)\\
&= \sum_a \tau^\mM \big(  w^* T_a  P T_a w \big) - \sum_a \tau^\mM \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \sum_a \tau_\infty \big( w^* T_a  ( P - w w^*) T_a w \big)\\
&\geq 1 - \eps -  \tau_\infty\Big(\big( P - w w^*\big)\Big(\sum_a  T_a w w^* T_a\Big)\Big)\\ 
&\geq 1- \eps- \tau_\infty\big( P- w w^*\big)\;,
\end{align*}
where the third line uses that $T_aPT_a=T_a$, $\sum_a T_a = \Id_\mN$ and the definition of $\eps$ for the first term, and for the second the fact that for $A\in\mM$, $\tau^\mM(A)=\tau_\infty(A)$ by definition of $\tau_\infty$ and the identification of $\mM$ with a ``corner'' in $\mM_\infty$, the fourth line uses cyclicity of the trace for the second, and the last uses $\|ww^*\|,\|\sum_a T_a\|_\infty\leq 1$. By assumption, 
\begin{align}
\tau_\infty\big( P- w w^*\big) \,\leq\, \eps\, \tau_\infty(P)\,\leq \frac{\eps}{1-\eps}\;,\label{eq:bound_on_P-ww*_in_trace_infty}
\end{align}
where the last inequality is because by definition, $\tau^N(P)=1$, thus
\[1-\eps \,\leq\, \tau^\mN(ww^*) \,=\, \frac{\tau_\infty(ww^*)}{\tau_\infty(P)}\,=\, \frac{\tau_\infty(w^*w)}{\tau_\infty(P)}  \,\leq\, \frac{1}{\tau_\infty(P)}\]
since $\tau_\infty(w^* w) = \tau^\mM(w^* w)$ and $w^*w\leq I_\mM$. Overall, 
\begin{equation}\label{eq:square_greater_1-3eps}
    \sum_a \tau^\mM \big( \tilde{Q}_a^2 \big) \,\geq\, 1-\eps-\frac{\eps}{1-\eps}\,\geq\, 1-3\eps\;.
\end{equation}
To conclude we apply~\cite[Theorem 1.2]{de2021orthogonalization} to obtain a projective measurement $\{Q_a\}$ on $\mM$ such that 
\begin{equation*}
\sum_a \big\|{Q}_a - \tilde{Q}_a \big\|^2_\tau \,=\, 27\eps\;.
\end{equation*}
Finally,
\begin{align*}
\sum_a \big\|{Q}_a - w^*{T}_a w\big\|^2_\tau &= \sum_a \Big\|{Q}_a - \tilde{Q}_a  + \frac{1}{|\mA|}\big(I_\mM - w^* w\big) \Big\|^2_2\\
&\leq  \sum_a 2\big\|{Q}_a - \tilde{Q}_a\big\|_\tau^2  + 2\frac{1}{|\mA|}\big\|I_\mM - w^* w\big\|_\tau^2 \\
&\leq 54 \eps + 2 \tau^\mM( (I_\mM - w^* w)^2 ) \\
&\leq 54 \eps + 2 \tau^\mM(I_\mM - w^* w ) \\
&\leq 56 \eps\;,
\end{align*}
where the second line is by the triangle inequality, the fourth line is due to the fact that $I_\mM - w^* w$ is positive and has operator norm at most $1$, and the last line is by $\tau^\mM(I_\mM - w^* w ) \leq \eps$.
\end{proof}

\begin{claim}\label{claim:L^1_dist_on_reps)_implies_L^infty_dist}
    Let $\phi_1\colon G\to \mU(\mM)$ and $\phi_2\colon G\to \mU(\mN)$ be representations of a finite group $G\cong \langle S\colon R\rangle$, where $\mN=PM_\infty P$  with  $\tau_\infty(P)<\infty$, and equipped with the appropriate trace. Let $w\in P\mM_\infty \Id_\mM$ be a partial isometry from $\mM$ to $\mN$, and assume both
    \[
    \Expectation_{x\in G}\Vert \phi_1(x)-w^*\phi_2(x)w\Vert_\tau ^2\leq \eps\quad \textrm{and}\quad \max\{\tau^\mM(\Id_\mM-w^*w),\tau^\mN(P-ww^*)\}\leq \eps.
    \]
    Then 
    \[
        \forall x\in G\ \colon\ \Vert \phi_1(x)-w^*\phi_2(x)w\Vert_\tau ^2\leq 12\eps.
    \]
    Namely, an $L^1$ bound on the distance between two representations implies an $L^\infty$ bound on their distance. In particular, the average distance along every generating set $S$ of $G$ is bounded by the expected distance over all of $G$.
\end{claim}
\begin{proof}
    The following idea is standard, and can be traced back, for example, to \cite{blum1990self}. As before, we can assume $\eps<\frac{1}{2}$, otherwise the claim is trivial. For every $x,y\in G$, we have 
    \begin{align}
      \Vert \phi_1(x)-w^*\phi_2(x)w\Vert_\tau&=\Vert \phi_1(y)\phi_1(y^{-1}x)-w^*\phi_2(y)\phi_2(y^{-1}x)w\Vert_\tau \notag\\
    &=\Vert \phi_1(y)\phi_1(y^{-1}x)-\phi_1(y)w^*\phi_2(y^{-1}x)w+\phi_1(y)w^*\phi_2(y^{-1}x)w \notag\\ 
    &-w^*\phi_2(y)ww^*\phi_2(y^{-1}x)w+w^*\phi_2(y)ww^*\phi_2(y^{-1}x)w-w^*\phi_2(y)\phi_2(y^{-1}x)w\Vert_\tau\notag \\
    &\leq \Vert \phi_1(y^{-1}x)-w^*\phi_2(y^{-1}x)w\Vert_\tau+\Vert \phi_1(y)-w^*\phi_2(y)w\Vert_\tau\notag\\
    &+\Vert w^*\phi_2(y)(ww^*-P)\phi_2(y^{-1}x)w\Vert_{\tau}.\label{eq:lower_bound_stab_standard_pres}
\end{align}
Now, the third summand in the last line of \eqref{eq:lower_bound_stab_standard_pres} can be bounded using  the  methods which appeared in the proof of Lemma \ref{lem:pull-back}:
\begin{align}
    \Vert w^*\phi_2(y^{-1}x)(ww^*-P)\phi_2(y)w\Vert_{\tau}^2&=\tau^\mM(w^*\phi_2^{-1}(y^{-1}x)(ww^*-P)\phi_2^{-1}(y)ww^*\phi_2(y)(ww^*-P)\phi_2(y^{-1}x)w)\notag\\
    &\leq \tau^\infty(w^*\phi_2^{-1}(y^{-1}x)(ww^*-P)\phi_2^{-1}(y)ww^*\phi_2(y)(ww^*-P)\phi_2(y^{-1}x)w)\notag\\
     &\leq \tau^\infty((ww^*-P)\phi_2^{-1}(y)ww^*\phi_2(y)(ww^*-P)\phi_2(y^{-1}x)ww^*\phi_2^{-1}(y^{-1}x))\notag\\
     &\leq \tau^\infty((ww^*-P)\phi_2^{-1}(y)ww^*\phi_2(y)(ww^*-P))\notag\\
     &\leq \tau^\infty(\underbrace{(ww^*-P)^2}_{P-ww^*}\phi_2^{-1}(y)ww^*\phi_2(y))\notag\\
     &\leq \frac{\eps}{1-\eps}\ \ ,\label{eq:lower_bound_stab_standard_pres1}
\end{align}
where we use the cyclicity of the trace, the fact $\mM$ is a corner and the evaluated operator is positive, the fact $\phi_2(y^{-1}x)ww^*\phi_2^{-1}(y^{-1}x)$ and $\phi_2^{-1}(y)ww^*\phi_2(y)$ are  contractions, as well as \eqref{eq:bound_on_P-ww*_in_trace_infty}.
Combining \eqref{eq:lower_bound_stab_standard_pres} and \eqref{eq:lower_bound_stab_standard_pres1}, we deduce that 
\begin{align*}
    \Vert \phi_1(x)-w^*\phi_2(x)w\Vert_\tau^2&=\Expectation_{y\in G}\Vert \phi_1(x)-w^*\phi_2(x)w\Vert_\tau^2\\
    &\leq 3\Expectation_{y\in G} [\Vert \phi_1(y^{-1}x)-w^*\phi_2(y^{-1}x)w\Vert_\tau^2+\Vert \phi_1(y)-w^*\phi_2(y)w\Vert_\tau^2\\
    &+\Vert w^*\phi_2(y)(ww^*-P)\phi_2(y^{-1}x)w\Vert_\tau^2 ]\\
    &\leq 3(\eps +\eps +\frac{\eps}{1-\eps})\leq 12\eps.
\end{align*}
\end{proof}

@@

As one would naturally expect, the dependence of $\delta_k$ on $k$ depends on both the diameter of $\Z_2^k$ for the presentation~\eqref{eq:z2-efficient}, and its \emph{Dehn function} --- namely, the minimal volume of a Van Kampen diagram with perimeters of length $3$. By expressing each element of $\Z_2^k$ as a product of generators in the natural way, applying Corollary~\ref{cor:lin-test} it is possible to show that $\delta_k=O(k^2\eps)$ in Lemma~\ref{lem:eff-z2}. With more work one can get $\delta_k=O(k\eps)$, see~\cite[Theorem 3.2]{chao2017overlapping}. This turns up to be  tight:
\begin{claim}
    Let $\mu_R$ be the uniform distribution over the commutation relations in the presentation $\langle S\colon R\rangle$ from \eqref{eq:z2-efficient}, and $\mu_S$ the marginal distribution from $\mu_R$, which is the uniform one on $S=\{x_1,...,x_k\}$. There is a $(\nicefrac{1}{\binom{k}{2}},\mu_R)$-almost homomorphism of $\langle S\colon R\rangle$ which is at least $\nicefrac{c}{k}$ far from any homomorphism. Namely, the modulus of stability of \eqref{eq:z2-efficient} is $\Omega(k\eps)$.
\end{claim}

@@

Now, all of the $A_i$'s are order $2$ unitaries, thus satisfying the order $2$ relations in \eqref{eq:z2-efficient} with probability $1$. Furthermore, 
\[
\Expectation_{i\neq j\in [k]}\left[\Vert A_iA_jA_iA_j-\Id_\mM\Vert_\tau^2\right]=\frac{1}{\binom{k}{2}}\Vert A_1A_2A_1A_2-\Id_\mM\Vert_\tau^2=\nicefrac{2}{\binom{k}{2}}.
\]
Thus, the map $x_i\mapsto A_i$ is a $(\nicefrac{1}{\binom{k}{2}},\mu_R)$-almost homomorphism of $\Z_2^k$ with respect to \eqref{eq:z2-efficient}. 
On the other hand, let $\{B_i\}_{i=1}^k$ be order $2$ unitaries which pairwise commute in $\mN=P\mM_\infty P $, and assume  there is a partial isometry $w\in P\mM_\infty\Id_\mM$ such that 
\[
\Expectation_{i\in [k]} \Vert A_i-w^* B_iw\Vert_\tau^2\ ,\  \tau^\mM(\Id_\mM-w^*w)\leq  \eps.
\]
This is a slightly weaker condition than for the $B_i$'s to be $(\eps,\mu_S)$-close to the $A_i$'s, as in Definition \ref{def:close}. 
Recall that $\mM_\infty$ acts on the Hilbert space $\complex^{\field^{k+1}}\otimes \complex^\Z$. Let $\{e_v\otimes e_t\mid v\in \field^{k+1},t\in \Z\}$ be the standard basis of this Hilbert space.

Then, by Lemma \ref{lem:pull-back} and Claim \ref{claim:L^1_dist_on_reps)_implies_L^infty_dist}, there are order $2$ unitaries $\{C_i\}_{i=1}^k$ in $\mM$, which pairwise commute and satisfy
\begin{equation*}
    \forall i\in [k]\ \colon \ \ \Vert C_i-w^*B_iw\Vert_\tau^2\leq 12\eps.
\end{equation*}
By the same argument as in \eqref{eq:square_greater_1-3eps}, we have 
\[
\tau^\mM((w^*B_iw)^2)\geq 1-3\eps,
\]
and thus
\[
\begin{split}
    12\eps&\geq \Vert C_i-w^*B_iw\Vert_\tau^2\\
    &=\underbrace{\tau^\mM(C_i^2)}_{=1}+\tau^\mM((w^*B_iw)^2)-2\tau^\mM(C_iw^*B_iw)\\
    &\geq 2-3\eps-2\tau^\mM(C_iw^*B_iw),
\end{split}
\]
which implies
\begin{equation}\label{eq:C_close_to_B}
    \tau^\mM(C_iw^*B_iw)\geq 1-8\eps.
\end{equation}
But, 
\begin{align}
    \sqrt{2}&=\Vert A_1A_2A_1A_2-\Id_\mM\Vert_\tau\notag\\
    &=\Vert A_1A_2A_1A_2-C_1C_2C_1C_2\Vert_\tau\notag\\
    &\leq 2\Vert A_1-C_1\Vert_\tau +2\Vert A_2-C_2\Vert_\tau,\label{eq:A_close_to_C}
\end{align}
where the inequality uses the triangle inequality and the fact $\{A_i\},\{C_i\}$ are unitaries, and the first equality is a by checking that $(A_1A_2)^2$ has no fixed points when acting on the set $\field^{k+1}$. Thus, for $i=1,2$, we have
\[
\begin{split}
    \tau^\mM(A_iw^*B_iw)&=\tau^\mM(A_iC_i)+\tau^\mM(A_i(w^*B_iw-C_i))
\end{split}
\]
Hence,
\begin{align*}
    k\eps&\geq \sum_{i=1}^k \Vert A_i-w^*B_iw\Vert_\tau ^2\\
&\geq \Vert A_1-w^*B_1w\Vert_\tau ^2+\Vert A_2-w^*B_2w\Vert_\tau ^2\\
&\geq4-2\eps-2\tau^
\mM(A_1w^*B_1w)-2\tau^\mM(A_2w^*B_2w)\\
&=4-2\eps-2\tau^\mM(C_1w^*B_1w)-2\tau^\mM((A_1-C_1)w^*B_1w)
\end{align*}
\end{proof}
